---
title: genepi.utils
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{css, include = FALSE, echo=FALSE}
p.caption {
  font-size: 0.7em;
}
```

<!-- badges: start -->
<!-- badges: end -->

The `genepi.utils` package is a collection of utility functions for working with genetic epidemiology data. For common use cases please see the [vignettes](https://nicksunderland.github.io/genepi.utils/).

## Installation

You can install the development version of genepi.utils like so:

```{r install, eval=FALSE}
# install.packages("devtools")
devtools::install_github("nicksunderland/genepi.utils")
```

The package relies on the parallel processing capabilities of the [`data.table`](https://rdatatable.gitlab.io/data.table/), [`furrr`](https://furrr.futureverse.org) and [`fst`](https://www.fstpackage.org) packages. These can be installed like so: 

```{r install_deps, eval=FALSE}
install.packages("data.table")
install.packages("furrr")
install.packages("fst")
```

**Important**: on mac it can be more challenging to enable OpenMP parallel processing as the *clang* compiler does not include an OpenMP runtime as standard. I recommend following the instructions on the `data.table` GitHub [link](https://github.com/Rdatatable/data.table/wiki/Installation). Key for successful installation on my Macbook M2 Max was creating a Makevars file in the root directory `~.R/Makevars`, which is a simple text file, containing the compilation flags below, and then re-installing from source. 

```{bash, eval=FALSE}
LDFLAGS += -L/opt/homebrew/opt/libomp/lib -lomp
CPPFLAGS += -I/opt/homebrew/opt/libomp/include -Xclang -fopenmp
```

Re-install packages from source.
```{r reinstall, eval=FALSE}
remove.packages("data.table")
remove.packages("furrr")
remove.packages("fst")

install.packages("https://cran.r-project.org/src/contrib/data.table_1.14.8.tar.gz", type="source", repos=NULL)
install.packages("https://cran.r-project.org/src/contrib/furrr_0.3.1.tar.gz", type="source", repos=NULL)
install.packages("https://cran.r-project.org/src/contrib/fst_0.9.8.tar.gz", type="source", repos=NULL)
```


## Download the dbSNP data repository

For RSID mapping you will need a copy of the dbSNP `.fst` file directory. Until this is hosted somewhere you will need to ask nicholas.sunderland@bristol.ac.uk for the data files. If you are working at the Bristol IEU the package should work on BC4. If not, ask for the location of the dbSNP `.fst` file directory on the HPC. If you use a custom location for the dbSNP directory you will need to set this in the package - this only needs to be done once per package install.

```{r set_dir, eval=FALSE}
set_dbsnp_directory("/path_to_directory/dbsnp")
```
