---
title: "Standardise GWAS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Standardise GWAS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here is a simple example of using the `standardise_gwas()` function to standardise your GWAS input ready for use with other features in this package. 

```{r setup}
library(genepi.utils)

gwas <- data.table::fread(system.file("extdata", "example2_gwas_sumstats.tsv", package="genepi.utils"))

gwas
```

## Column mapping

We provide a number of column mapping options for standard GWAS formats through the `ColumnMapping()` class. The available formats can be queried like so: 

```{r mapping}
# column mapping object
map_obj <- ColumnMapping()

# available formats
input_formats(map_obj)
```

## Standardise GWAS

A GWAS can then be standardised like so:

```{r std}
gwas <- standardise_gwas(gwas, input_format="ns_map", build="GRCh37", populate_rsid=FALSE)[]

gwas
```

## Custom mapping

The provided mapping formats can be inspected with `get_map()`. The `input_format` input of `standardise_gwas()` can either be one of the pre-defined mappings (as shown above), or you can provide your own custom mapping in the form of a named list. In this case the list should contain a minimum of the names SNP, CHR, BP, EA, OA, BETA, SE, P, and EAF.

```{r format}
map_obj <- ColumnMapping()

standard_mapping <- get_map(map_obj, input_format="ns_map")

standard_mapping
```


```{r custom_map, eval=FALSE}
custom_map <- list(SNP  = "snp_marker", 
                   CHR  = "chrom", 
                   BP   = "position", 
                   EA   = "effect_allele", 
                   OA   = "alt_allele", 
                   BETA = "beta", 
                   SE   = "std_error", 
                   P    = "pval", 
                   EAF  = "frq")

# not run
gwas <- standardise_gwas(gwas, input_format=custom_map, build="GRCh37", populate_rsid=FALSE)[]
```
